Forecasting the 2024 U.S. Presidential Election: Kamala Harris' Projected Victory

### 2024-11-04 paper/paper.qmd
committing changes to paper.qmd for formatting purposes

#### Numeric Grade
Numeric Grade (**'numeric_grade'**) assesses the quality of each poll, as rated by Project 538 [@FiveThirtyEight], based on factors like methodology, transparency, and historical accuracy. In our paper, this variable allows us to weigh the reliability of different polls, ensuring that those with higher ratings contribute more significantly to our voter support predictions. 
```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false
#| label: fig-numgrade-distribution
#| fig-cap: "Density plot illustrating the distribution of numeric grades assigned to pollsters within the polling data. Highlights the frequency of numeric grades, indicating key concentrations that suggest variability in pollster quality. "

@@

```
@fig-numgrade-distribution illustrates the distribution of numeric grades in our analysis, which ranges from 2.5 to 3.0. Higher grades indicate more methodologically rigorous and transparent polls, enhancing the reliability of our model's predictions. The distribution shows several peaks around grades 2.7, 2.8, and 3.0, suggesting that certain pollsters consistently achieve higher ratings due to better methodologies or transparency. Conversely, valleys in the distribution may reflect a scarcity of polls from lower-rated sources. By incorporating numeric grade, we enhance prediction accuracy by prioritizing data from more reliable sources, which is particularly crucial in closely contested states where polling accuracy is essential for reliable forecasts.

@@

# Distribution of start_date with x-axis labels every 15 days
ggplot(clean_poll_data, aes(x = start_date)) +
  geom_histogram(bins = 30, fill = "lightblue", color = "black") +
  labs(
       x = "Start Date",
       y = "Count") +
  scale_x_date(date_breaks = "15 days", date_labels = "%b %d") +  # Adjust x-axis labels
  theme_minimal()
```
@fig-start-date-distribution displays the distribution of polling start dates from late July to mid-October 2024, showing when polling efforts began across key battleground states. The histogram reveals peaks in polling activity, particularly in mid-August, early September, and late September, indicating periods of heightened interest or campaign events that likely prompted increased polling. This distribution provides insights into the timing and intensity of polling efforts throughout the campaign season. By including only polls initiated after Harris’s campaign launch, we capture voter preferences in a context where both major candidates are actively campaigning, thus making the polling data more relevant for direct comparisons. This filtered timeline ensures our model focuses on data that reflects the competitive dynamics between Harris and Trump, enhancing the validity of our predictive analysis as we approach Election Day.

>>>>

#### Numeric Grade
Numeric Grade (**'numeric_grade'**) assesses the quality of each poll, as rated by Project 538 [@FiveThirtyEight], based on factors like methodology, transparency, and historical accuracy. In our paper, this variable allows us to weigh the reliability of different polls, ensuring that those with higher ratings contribute more significantly to our voter support predictions. @fig-numgrade-distribution illustrates the distribution of numeric grades in our analysis, which ranges from 2.5 to 3.0. Higher grades indicate more methodologically rigorous and transparent polls, enhancing the reliability of our model's predictions. The distribution shows several peaks around grades 2.7, 2.8, and 3.0, suggesting that certain pollsters consistently achieve higher ratings due to better methodologies or transparency. Conversely, valleys in the distribution may reflect a scarcity of polls from lower-rated sources. By incorporating numeric grade, we enhance prediction accuracy by prioritizing data from more reliable sources, which is particularly crucial in closely contested states where polling accuracy is essential for reliable forecasts.
```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false
#| label: fig-numgrade-distribution
#| fig-cap: "Density plot illustrating the distribution of numeric grades assigned to pollsters within the polling data. Highlights the frequency of numeric grades, indicating key concentrations that suggest variability in pollster quality. "

@@

@@

# Distribution of start_date with x-axis labels every 15 days
ggplot(clean_poll_data, aes(x = start_date)) +
  geom_histogram(bins = 30, fill = "lightblue", color = "black") +
  labs(
       x = "Start Date",
       y = "Count") +
  scale_x_date(date_breaks = "15 days", date_labels = "%b %d") +  # Adjust x-axis labels
  theme_minimal()
```
@fig-start-date-distribution illustrates the distribution of polling start dates from late July to mid-October 2024, highlighting when polling efforts commenced in key battleground states. The histogram shows peaks in activity during mid-August, early September, and late September, suggesting increased polling during significant campaign events. By including only polls initiated after Harris's campaign launch, we capture voter preferences in a context where both major candidates are actively campaigning. This approach ensures our model focuses on data that accurately reflects the competitive dynamics between Harris and Trump, enhancing the validity of our predictive analysis as Election Day approaches.
